Code walkthrough

### 2022-10-08 src/explore_aghq/walkthrough.Rmd
aghq::sample_margial step-by-step

#' Replacing the theta, theta1, theta2, ... names with the names from TMB
if (is.null(thetanames)) {
  thetanames <- colnames(distinctthetas)
} else {
  colnames(modesandhessians) <- thetanames
  colnames(quad$normalized_posterior$nodesandweights)[grep('theta',colnames(quad$normalized_posterior$nodesandweights))] <- thetanames
}

@@

# `aghq::sample_marginal`

@@

>>>>

#' Replacing the theta, theta1, theta2, ... names with the names from TMB
if (is.null(thetanames)) {
  thetanames <- colnames(distinctthetas)
} else {
  colnames(modesandhessians) <- thetanames
  colnames(quad$normalized_posterior$nodesandweights)[grep('theta', colnames(quad$normalized_posterior$nodesandweights))] <- thetanames
}

@@

# `aghq::sample_marginal.marginallaplace`

@@

# `aghq::sample_marginal.aghq`

Let's have a look at what that call to `sample_marginal.aghq` is doing:

```{r}
class(quad)

out <- list()

#' If the quadrature object somehow doesn't contain any information about the 
if (is.null(quad$marginals)) return(out)

str(quad$marginals)
```

As usual, we're going to start by doing the computation for the first element of `theta` explicitly:

```{r}
i <- 1
```

We already have information about each marginal, contained in:

```{r}
quad$marginals[[i]]
```

So we can use the function `aghq::compute_quantiles` to, my guess, use the inverse CDF method to sample from the marginal, given input of some $\mathcal{U}(0, 1)$ draws:

```{r}
unif <- stats::runif(M)
unif

compute_quantiles(quad$marginals[[i]], unif, interpolation = interpolation, transformation = transformation)
```

```{r}
for (i in 1:length(quad$marginals)) {
  out[[i]] <- unname(compute_quantiles(quad$marginals[[i]], stats::runif(M), interpolation = interpolation, transformation = transformation))
}

#' The function doesn't name these yet, but in future it could (and also return a data.frame) via
d <- length(quad$optresults$mode)
thetanames <- colnames(quad$normalized_posterior$nodesandweights)[1:d]
names(out) <- thetanames
as.data.frame(out)
```

# `aghq::compute_quantiles.aghq`

Does the `compute_quantiles` function work as we expect it to?

```{r}
class(quad$marginals[[i]])
```

The PDF and CDF are computed by `compute_pdf_and_cdf` (as the name suggests, and another layer deep that we can look).
Code comments note that transformations are not applied at this point, and that it's better to transform the quantiles later:

```{r}
q <- stats::runif(M)

pdfandcdf <- compute_pdf_and_cdf(quad$marginals[[i]], interpolation = interpolation)

ggplot(pdf_and_cdf, aes(x = theta, y = cdf)) +
  geom_line() +
  theme_minimal()

out <- numeric(length(q))
increasing <- TRUE

validate_transformation(transformation)
transformation <- make_transformation(transformation)
```

For each sample, take the maximum `theta` which still has CDF value less than the quantile desired:

```{r}
for (i in 1:length(q)) out[i] <- pdfandcdf$theta[max(which(pdfandcdf$cdf < q[i]))]

#' Have to check if it's increasing or decreasing so can reverse order if necessary
increasing <- transformation$fromtheta(min(out)) <= transformation$fromtheta(max(out))
for (i in 1:length(out)) out[i] <- transformation$fromtheta(out[i])

if (!increasing) out <- rev(out)
names(out) <- paste0(as.character(100 * q), "%")
out
```

# `compute_pdf_and_cdf`

Going to leave this as a to-do.
